%!TEX root = ../thesis.tex
\null\vfill

\section*{Acknowledgements}
	
	As I start writing these lines intended to thank my \emph{Doktorvater} Kurt Mehlhorn, I struggle to find the words capturing the depth of my gratitude towards him. Kurt, if you read this, please know that I will never forget the role you played in my life.

	To me (and to many others) you are, and always will be, a true role model worthy of admiration. You are a just and caring person, an enormous scientist, an unparalleled teacher, a wise mentor and an exemplary leader. I admire you for your humble demeanour, your perfect diligence and your unwavering purpose. Your down-to-earth attitude and your seemingly untiring willingness to help others wherever you can uplifts and inspires everyone around you.

	I am grateful for the privilege of being your student. I feel extremely humbled by your unending support and constant encouragement. I am grateful that you bravely trusted me to find myself through my own projects, even if they were risky and out of the ordinary. At times, I didn't feel up to the task myself. Frequently I struggled. Nevertheless, whatever the problem seemed to be, you never failed to help me find the missing pieces. In truth, your generous advice quietly enabled me every step of the way. I will carry with me the numerous lessons you taught me about science, about people and about life. For this I will always be indebted to you. In fact, I hope that one day, I will be able to do for someone, what you have done for me.\bigskip
	
	In addition to Kurt, there are many others that deserve my gratitude. Without them this thesis would not have been possible either. This is particularly true for my many outstanding co-authors. True to the nature of modern science, most of the results presented in this thesis are the product of various collaborations and many hours of lively discussions amongst colleagues. I would be remiss If I didn't mention Adrian Neumann and Matthias F\"ugger who shared my ideas and supported me actively during the earlier and the later stages of my doctorate respectively. I estimate that roughly two thirds of my sanity have been preserved by interacting with those two individuals and of course, by procrastinating with my office mates.\bigskip

	My final thanks are dedicated to my dear Sarah. I know that the ``last hurrah'' of completing this thesis was as least as hard on you as it was on me. Thank you for being patient and for overlooking, without complaint, the many late-night sessions it required. Thank you also for always being convinced that my stuff was solid, and that I was solid, and that I would pull through in the end no matter what.


	% Dear Kurt, frankly, I do not know how you do what you do. It is a mystery to me how you do it day in day. What I do know is that I have yet to meet someone who does not admire you in some way or another. 

	% I admire you for your calm and humble demeanor.
	% I admire you for your diligence and unwavering purpose.
	% I admire you for your gift of explanation and your unending patience with students. 
	% I admire you for always asking the questions that need to be asked, whether they are pleasant or not.
	% I admire you for always having the missing piece sitting somewhere in you back pocket.
	% I admire you for showing up at work smiling with a gigantic band-aid plaster right across your face telling everyone what happed with a laugh.
	% I admire you laid-back attitude and positive outlook, always trusting that things will work out in the end.

	% You groomed us to be sound professionals and made working with you an interesting and memorable experience. I will always be grateful to you for your support and kindness.

	% Not only have you been a fantastic mentor to me, but you have taught me how to mentor other people. Thank you for being such a great role model.

	% You have been an exemplary and visionary mentor, a great leader who has dedicated his life to the service of humanity. I appreciate and treasure everything you have taught me.

	% changed the trajectory of my life

	% Your encouragement and advice has led me to places I never thought I would go. 

	% I hope one day I will be able to do for someone what you have done for me.

	\vfill